---
title: "Fisher"
output: html_document
date: "2023-02-22"
---

This stuff is modified from  [RSF and SSF analysis of fisher. "S. Muff, J. Signer, J. Fieberg"](https://conservancy.umn.edu/bitstream/handle/11299/204737/fisher_rsf_and_ssf.R?sequence=26&isAllowed=y)

and

https://conservancy.umn.edu/bitstream/handle/11299/218272/AppB_SSF_examples.html?sequence=26

The data are from - LaPoint, S., Gallery, P., Wikelski, M. and Kays, R. (2013). Animal behavior, cost-based corridor models, and real corridors. Landscape Ecology, 28, 1615-1630.


```{r}
library(tidyverse)
library(raster)
library(rgdal)
library(amt)
library(cmdstanr)
```


#' Now lets start by reading in the simplified fisher data
```{r}
dat <- read_csv("fisher_data.csv")
```

Include sex of each animal and create tracks with an appropriate coordinate reference system using the amt package

```{r}
dat_all <- dat %>% nest(-id) 
dat_all$sex <- c("f", "f", "f", "m", "m", "m")
dat_all <- dat_all %>% 
  mutate(trk = map(data, function(d) {
    make_track(d, x, y, t, crs = sp::CRS("+init=epsg:4326")) %>% 
      transform_coords(sp::CRS("+init=epsg:5070"))
  }))
```

Summarize sampling rates, 
```{r}
dat_all %>% mutate(sr = lapply(trk, summarize_sampling_rate)) %>% 
  select(id, sr) %>% unnest

#' 10 minutes seems to appropriate for all animals.
#' Resample the track to 10 minutes with a tolerance of 2 minutes.

dat1 <- dat_all %>% mutate(dat_clean = map(trk, ~ {
  .x %>% track_resample(rate = minutes(10), tolerance = seconds(120))
  }))

```

Read in the landuse raster and reclassify to two categories (wet forests and other).
```{r}
landuse <- amt_fisher_covar$landuse
elevation <- amt_fisher_covar$elevation
popden <- amt_fisher_covar$popden
reclass_landuse <- function(x) {
  fct_collapse(factor(x),
               forest = c("30","40","50","60", "70","80", "90","100"),
               grass = c("120", "140"),
               wet = c("160"))
}

```


## Step-Selection Function (SSF)

Data development for step-selection function

First we need to prepare the data. We need to pair each observed point with 10 random points and extract the covariate value at the end point of each step.

```{r}

#+ warning = FALSE
dat_ssf <- dat1 %>% 
  mutate(stps = map(dat_clean, ~ .x %>% 
                      steps_by_burst() %>% 
                      random_steps() %>% 
                      extract_covariates(wet_forests) %>%
                      extract_covariates(elevation) %>%
                      extract_covariates(popden))) %>% 
  select(id, stps) %>% unnest() %>% 
  mutate(
    y = as.numeric(case_),
    id = as.numeric(factor(id)), 
    step_id = paste0(id, step_id_, sep = "-"))

```


```{r}

tmp = which(is.na(dat_ssf$elevation))
dat_ssf = dat_ssf[-tmp,]
tmp = which(is.na(dat_ssf$popden))
dat_ssf = dat_ssf[-tmp,]
dat_ssf$id = as.numeric(as.factor(dat_ssf$id))

x = matrix(NA, nrow(dat_ssf), 3)
x[,1] = dat_ssf$layer
x[,2] = scale(dat_ssf$elevation)
x[,3] = scale(dat_ssf$popden)

tmp = as.numeric(as.factor(dat_ssf$step_id))
tmp1 = numeric(length(tmp))

tmp1[1] = 1
for(i in 2:length(tmp)){
  tmp1[i] = tmp1[i-1]
  if(tmp[i] != tmp[i-1]){
    tmp1[i] = tmp1[i]+1
  }
}
```


```{r}
stan_dat <- list(N = nrow(dat_ssf), 
                 I = length(unique(dat_ssf$step_id)), 
                 J = length(unique(dat_ssf$id)), 
                 y = dat_ssf$y, 
                 x = x, 
                 K = ncol(x),
                 stepid = tmp1, 
                 indid = dat_ssf$id)
```


```{r}
mod = cmdstan_model('PoissonSSF.stan')
```

```{r}
init_fun <- function() list(a_re = rep(0, length(unique(dat_ssf$step_id))),
                               beta_p = matrix(0,5,3))
```


```{r}
fit = mod$sample(
  data = stan_dat,
  init = init_fun,
  chains = 4,
  parallel_chains = 4,
  iter_warmup = 1000,
  iter_sampling = 1000,
  thin = 1)
```

```{r}
pars = c("beta_p", "beta")
fit_summary = fit$summary(pars)
```

